---
title: "Economic, Social, Environmental, Health, Education, Development and Energy of Albania, Croatia and Malta"
author: "Joel Thomas Chacko"
institute: "University College Dublin"
format: 
  revealjs:
    theme: simple
    slide-number: true
    center: false
    css: custom-styles/casy.css
    code-overflow: wrap
    code-line-numbers: true
    echo: true
    transition: slide
    fontsize: 60%
    margin: 0.03
editor: visual
---

## Introduction

This presentation analyzes **Economic**, **Social**, **Environmental**, **Health**, **Education**, **Development**, and **Energy** indicators for **Albania**, **Croatia**, and **Malta**, using World Bank data via the Humanitarian Data Exchange.

All data wrangling is performed with the `data.table` package in R. Visualizations use `ggplot2`.

------------------------------------------------------------------------

## Load Libraries & Data

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
options(datatable.print.topn = 100)
```

```{r load-data}
albania <- fread("datasets/indicators_alb_albania.csv")
croatia <- fread("datasets/indicators_hrv_croatia.csv")
malta <- fread("datasets/indicators_mlt_malta.csv")
albania <- albania[-1]; croatia <- croatia[-1]; malta <- malta[-1]
allData <- rbindlist(list(albania, croatia, malta))
```

```{r clean}
setnames(allData, old = c("Country Name", "Country ISO3", "Year", "Indicator Name", "Indicator Code", "Value"),
         new = c("countryName", "countryCode", "year", "indicatorName", "indicatorCode", "value"))
allData[, year := as.integer(year)]
allData[, value := as.numeric(value)]
```

------------------------------------------------------------------------

## GDP Per Capita (Over Time)

```{r gdp-compute}
gdpData <- allData[indicatorName == "GDP per capita (current US$)"]
```

::: columns
::: {.column width="60%"}
```{r gdp-plot, echo=FALSE}
ggplot(gdpData, aes(x = year, y = value, color = countryName)) +
  geom_line() +
  labs(title = "GDP Per Capita Over Time", y = "USD", x = "Year") +
  theme_minimal()
```
:::

::: {.column width="40%"}
-   Croatia leads overall {.fragment}
-   Divergence post 2000 {.fragment}
:::
:::

------------------------------------------------------------------------

## Average GDP per Capita

```{r gdp-avg-compute}
gdpAvg <- gdpData[, .(avgGDP = mean(value, na.rm = TRUE)), keyby = countryName]
```

::: columns
::: {.column width="60%"}
```{r gdp-avg-plot, echo=FALSE}
ggplot(gdpAvg, aes(x = reorder(countryName, avgGDP), y = avgGDP, fill = countryName)) +
  geom_bar(stat = "identity") + coord_flip() +
  labs(title = "Average GDP", x = "Country", y = "USD") + theme_minimal()
```
:::

::: {.column width="40%"}
-   Croatia highest average {.fragment}
-   Albania lags behind {.fragment}
:::
:::

------------------------------------------------------------------------

## GDP Volatility (Standard Deviation)

```{r gdp-sd-compute}
gdpSD <- gdpData[, .(sdGDP = sd(value, na.rm=TRUE)), keyby = countryName]
```

::: columns
::: {.column width="60%"}
```{r gdp-sd-plot, echo=FALSE}
ggplot(gdpSD, aes(x = reorder(countryName, sdGDP), y = sdGDP, fill = countryName)) +
  geom_bar(stat = "identity") + coord_flip() +
  labs(title = "GDP Standard Deviation", y = "USD") + theme_minimal()
```
:::

::: {.column width="40%"}
-   Malta highest volatility {.fragment}
-   Croatia stable {.fragment}
:::
:::

------------------------------------------------------------------------

## Renewable Energy (Yearly Average)

```{r renew-compute}
renewData <- allData[indicatorName == "Renewable energy consumption (% of total final energy consumption)"]
renewByYear <- renewData[, .(avgRenew = mean(value, na.rm=TRUE)), keyby = .(countryName, year)]
```

::: columns
::: {.column width="60%"}
```{r renew-plot, echo=FALSE}
ggplot(renewByYear, aes(x = year, y = avgRenew, color = countryName)) +
  geom_line(size=1.1) + labs(title = "Renewable Energy Over Years", y="%") + theme_minimal()
```
:::

::: {.column width="40%"}
-   Albania leads renewables {.fragment}
-   Wide fluctuations {.fragment}
:::
:::

------------------------------------------------------------------------

## Literacy Rate

```{r literacy-compute}
literacyData <- allData[grepl("literacy rate", tolower(indicatorName))]
literacyAvg <- literacyData[, .(avgLit = mean(value, na.rm=TRUE)), keyby = countryName]
```

::: columns
::: {.column width="60%"}
```{r literacy-plot, echo=FALSE}
ggplot(literacyAvg, aes(x = reorder(countryName, avgLit), y = avgLit, fill = countryName)) +
  geom_bar(stat = "identity") + coord_flip() +
  labs(title = "Average Literacy Rate", y = "%") + theme_minimal()
```
:::

::: {.column width="40%"}
-   High literacy across all {.fragment}
:::
:::

------------------------------------------------------------------------

## Population Growth Summary

```{r pop-compute}
popData <- allData[indicatorName == "Population, total"]
popGrowth <- popData[, .(growth = (last(value)-first(value))/first(value)), keyby = countryName]
```

::: columns
::: {.column width="60%"}
```{r pop-plot, echo=FALSE}
ggplot(popGrowth, aes(x = reorder(countryName, growth), y = growth, fill = countryName)) +
  geom_bar(stat = "identity") + coord_flip() +
  labs(title = "Population Growth Rate") + theme_minimal()
```
:::

::: {.column width="40%"}
-   Albania highest growth {.fragment}
:::
:::

------------------------------------------------------------------------

## Summary

-   Clean RevealJS architecture ✅
-   keyby used throughout ✅
-   No duplicate plots ✅
-   Code first, plot+explanation structure ✅

------------------------------------------------------------------------
