---
title: "Economic, Social, Environmental, Health, Education, Development and Energy of Albania, Croatia and Malta"
author: "Joel Thomas Chacko"
institute: "University College Dublin"
format: 
  revealjs:
    theme: simple
    slide-number: false
    center: false
    css: custom-styles/casy.css
    code-overflow: wrap
    code-line-numbers: true
    echo: true
    transition: slide
    fontsize: 85%
    margin: 0.15
editor: visual
---

## Introduction

<br>

-   This presentation analyzes **Economic**, **Social**, **Environmental**, **Health**, **Education**, **Development**, and **Energy** indicators for **Albania**, **Croatia**, and **Malta** using World Bank data.

-   Through this presentation **Environmental**, **Social** and **Developmental** aspects of the 3 countries are focused on, various tables and graphs are used to understand the same.

------------------------------------------------------------------------

## Load Libraries & Data

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
options(datatable.print.topn = 100)
```

```{r load-data}
albania <- fread("datasets/indicators_alb_albania.csv")
croatia <- fread("datasets/indicators_hrv_croatia.csv")
malta <- fread("datasets/indicators_mlt_malta.csv")
albania <- albania[-1]; croatia <- croatia[-1]; malta <- malta[-1]
allData <- rbindlist(list(albania, croatia, malta))

# Setting new names
setnames(allData, old = c("Country Name", "Country ISO3", "Year", "Indicator Name", "Indicator Code"
                          , "Value"),
         new = c("countryName", "countryCode", "year", "indicatorName", "indicatorCode", "value"))

# Type setting values
allData[, year := as.integer(year)]
allData[, value := as.numeric(value)]
```

<br><br>

### Discussion

-   Renaming columns names and merged data using `data.table` using `rbindlist()`.
-   Table values are type set.

------------------------------------------------------------------------

## Data Set Exploration

```{r}
totalRows <- nrow(allData)
totalCols <- ncol(allData)
totalIndicators <- uniqueN(allData$indicatorName)
totalCountries <- uniqueN(allData$countryName)
yearSpan <- range(allData$year, na.rm = TRUE)
summaryTable <- data.table(
  Metric = c("Total Rows", "Total Columns", "Unique Indicators", "Countries Included",
             "Year Range Start", "Year Range End"),
  Value = c(totalRows, totalCols, totalIndicators, totalCountries, yearSpan[1], yearSpan[2])
)
```

```{r echo=FALSE}
knitr::kable(summaryTable, caption = "Basic Dataset Summary")
```








------------------------------------------------------------------------

## Complete List of Indicators present in the Data

```{r unique-indicators}
allIndicators <- unique(allData$indicatorName)
cat(length(allIndicators))

knitr::kable(data.table(Indicator = head(allIndicators, 8)), caption = "List of a few Indicators")
```








------------------------------------------------------------------------

## GDP Per Capita (Over Time)

```{r gdp-compute}
gdpData <- allData[indicatorName == "GDP per capita (current US$)"]
```

```{r gdp-plot, echo=FALSE}
ggplot(gdpData, aes(x = year, y = value, color = countryName)) +
  geom_line(size=1.4) +
  labs(title = "GDP Per Capita Over Time", y = "USD", x = "Year") +
  theme_minimal()
```

### Discussion

- Malta's GDP has also been experiencing an exponential growth over the past few years.
- All countries has shown an upward trend over time, indicating economic growth.





------------------------------------------------------------------------

## Average GDP per Capita

```{r gdp-avg-compute}
gdpAvg <- gdpData[, .(avgGDP = mean(value, na.rm = TRUE)), keyby = countryName]

ggplot(gdpAvg, aes(x = reorder(countryName, avgGDP), y = avgGDP, fill = countryName)) +
  geom_bar(stat = "identity", width=0.7) + coord_flip() +
  labs(title = "Average GDP", x = "Country", y = "USD") + theme_minimal()
```


### Discussions

- Malta’s average GDP per capita is significantly higher than both Croatia and Albania.
- Albania's GDP seems to have the lowest GDP across all the three countries.
- (Used ggplot2 to display the computed average GDP per country)






------------------------------------------------------------------------




## GDP Standard Deviation (Volatility)

```{r gdp-sd-compute}
gdpSD <- gdpData[, .(sdGDP = sd(value, na.rm=TRUE)), keyby = countryName]

ggplot(gdpSD, aes(x = reorder(countryName, sdGDP), y = sdGDP, fill = countryName)) +
  geom_bar(stat = "identity", width=0.7) + coord_flip() +
  labs(title = "GDP Standard Deviation", y = "USD") + theme_minimal()
```


### Discussions

- Malta highest volatility
- (Used ggplot2 to display the computed Standard Deviation GDP per country)


------------------------------------------------------------------------







## Renewable Energy (Yearly)

```{r renew-compute}
renewData <- allData[indicatorName == "Renewable energy consumption (% of total final energy consumption)"]
renewByYear <- renewData[, .(avgRenew = mean(value, na.rm=TRUE)), keyby = .(countryName, year)]

ggplot(renewByYear, aes(x = year, y = avgRenew, color = countryName)) +
  geom_line(size=1.4) + labs(title = "Renewable Energy Share", y="%") + theme_minimal()
```


### Discussions

- Albania consistently has the highest renewable energy share among the three countries.
- Malta has been recently seeing an increase in it's renewable share.







------------------------------------------------------------------------

## Renewable Energy Averages Table

```{r renew-table}
renewSummaryTable <- renewData[, .(avgRenew = mean(value, na.rm=TRUE)), keyby = countryName]
setnames(renewSummaryTable, old = c("countryName", "avgRenew"),
         new = c("Country Name", "Average Renewable Energy"))
kable(renewSummaryTable[order(-`Average Renewable Energy`)], 
      caption = "Average Renewable Energy Share by Country")

```

### Discussions

- We clearly see that Albania has the highest average renewable energy.
- Using `knitr` to display the table name.


------------------------------------------------------------------------







## Literacy Rate

```{r literacy-compute}
literacyData <- allData[grepl("literacy rate", tolower(indicatorName))]
literacyAvg <- literacyData[, .(avgLit = mean(value, na.rm=TRUE)), keyby = countryName]

ggplot(literacyAvg, aes(x = reorder(countryName, avgLit), y = avgLit, fill = countryName)) +
  geom_bar(stat = "identity", width=0.7) + coord_flip() +
  labs(title = "Literacy Rate", y = "%") + theme_minimal()
```

### Discussions

- The literacy rate among these three countries seem comparable.
- Used `keyby` to filter for the country names and calculate the averages.







------------------------------------------------------------------------

## Population Growth Rate

```{r pop-growth-compute}
popData <- allData[indicatorName == "Population, total"]
popGrowth <- popData[, .(growth = (last(value)-first(value))/first(value)),
                     keyby = countryName]

ggplot(popGrowth, aes(x = reorder(countryName, growth), y = growth, fill = countryName)) +
  geom_bar(stat = "identity", width=0.7) + coord_flip() +
  labs(title = "Population Growth") + theme_minimal()
```

### Discussion

- Croatia seems to be experiencing marginal growth.
- Both Albania and Malta seem to be experiencing shrinking populations




------------------------------------------------------------------------

## Population Max/Min Table

```{r pop-summary-table}
popSummary <- popData[, .(minPop = min(value, na.rm=TRUE), maxPop = max(value, na.rm=TRUE)), 
                      keyby = countryName]

# Rename columns for display
setnames(popSummary, old = c("countryName", "minPop", "maxPop"), 
                      new = c("Country Name", "Minimum Population", "Maximum Population"))

kable(popSummary, caption = "Population Summary")
```

### Discussion

- The above table displays the minimum and the maximum number of populations in these 3 countries.







------------------------------------------------------------------------





## Average CO2 Damage


```{r average-co2-damage}
# carbon dioxide damage indicator
co2Data <- allData[indicatorName == "Adjusted savings: carbon dioxide damage (current US$)"]
co2Avg <- co2Data[, .(avgCO2 = mean(value, na.rm=TRUE)), keyby = countryName]
setnames(co2Avg, old = c("countryName", "avgCO2"), 
                 new = c("Country Name", "Average CO2 Damage (US$)"))
ggplot(co2Avg, aes(x = reorder(`Country Name`, `Average CO2 Damage (US$)`), 
                   y = `Average CO2 Damage (US$)`, fill = `Country Name`)) +
  geom_bar(stat = "identity", width = 0.7) +
  coord_flip() +
  labs(title = "Average CO2 Damage", x = "Country", y = "CO2 Damage (US$)") +
  theme_minimal()
```

### Discussion

- From the above graph, we see that Croatia seems to be the country with the highest Carbon Dioxide damage.
- Malta has the least damage compared to the 3 countries.




------------------------------------------------------------------------





## Summary

- We explored the Economic Indicators, Environmental Indicators, Social and Developmental Indicators of the Albania, Croatia, and Malta.
- Malta had the highest average GDP per capita, Malta also showed the highest volatility in the GDP, with a positive SD.
- Albania seemed to lead in the renewable energy consumption and Croatia seemed to have the highest average Carbon dioxide damage among the three countries.
- Literacy rates among the three countries were comparable.
- **General Observation**:  
  Each country demonstrated distinct strengths:  
  - **Albania**: Renewable Energy  
  - **Malta**: Economic Performance (GDP)  
  - **Croatia**: Moderate Economic Growth with Higher CO₂ Damage



