---
title: "Economic, Social, Environmental, Health, Education, Development and Energy of Albania, Croatia and Malta"
author: "Joel Thomas Chacko"
institute: "University College Dublin"
format:
  beamer:
    theme: Madrid
    colortheme: seahorse
    fonttheme: professionalfonts
    include-in-header: custom-meta-data/custom-tex.tex
editor: visual
---

## Introduction

This presentation analyzes **Economic**, **Social**, **Environmental**, **Health**, **Education**, **Development**, and **Energy** indicators for **Albania**, **Croatia**, and **Malta**, using World Bank data via the Humanitarian Data Exchange.

All data wrangling is performed with the `data.table` package in R. Visualizations use `ggplot2`.

------------------------------------------------------------------------

## Data Preparation

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
options(datatable.print.topn = 100)
```

------------------------------------------------------------------------

## Load and Combine Datasets

```{r load-data}
albania <- fread("datasets/indicators_alb_albania.csv")
croatia <- fread("datasets/indicators_hrv_croatia.csv")
malta <- fread("datasets/indicators_mlt_malta.csv")

# Drop headers
albania <- albania[-1]
croatia <- croatia[-1]
malta <- malta[-1]

# Merge using functionality provided by data.table
allData <- rbindlist(list(albania, croatia, malta))
```

------------------------------------------------------------------------

## Clean Column Names and Types

```{r clean}
#| echo: true
setnames(allData, old = c("Country Name", "Country ISO3", "Year", "Indicator Name", "Indicator Code", "Value"),
                 new = c("countryName", "countryCode", "year", "indicatorName", "indicatorCode", "value"))

allData[, year := as.integer(year)]
allData[, value := as.numeric(value)]
```

------------------------------------------------------------------------

## Quick Exploration

```{r explore}
allData[, .N, by = .(countryName)]
allData[, .N, by = .(indicatorName)][1:10]
```

------------------------------------------------------------------------

## Sample Plot: Population Trend

```{r pop-plot}
popData <- allData[indicatorName == "Population, total"]
ggplot(popData, aes(x = year, y = value, color = countryName)) +
  geom_line(size = 1.2) +
  labs(title = "Population Growth", y = "Population", x = "Year") +
  theme_minimal()
```

------------------------------------------------------------------------

## GDP Per Capita

```{r}

gdpData <- allData[indicatorName == "GDP per capita (current US$)"]
ggplot(gdpData, aes(x = year, y = value, color = countryName)) +
  geom_line() +
  labs(title = "GD##P per Capita Over Time", y = "USD", x = "Year") +
  theme_minimal()

```

## Top 10 Countries by Average GDP Per Capita

```{r}

library(knitr)

gdpSubset <- allData[indicatorName == "GDP per capita (current US$)"]

topGDP <- gdpSubset[, .(avgGDP = mean(value, na.rm = TRUE)), keyby = countryName][
  order(-avgGDP)][1:10]

kable(topGDP, caption = "Top 10 Countries by Average GDP per Capita")


```

## Literacy Emissions

```{r}

literacyData <- allData[grepl("literacy rate", tolower(indicatorName))]
ggplot(literacyData, aes(x = year, y = value, color = countryName)) +
  geom_line() +
  labs(title = "Literacy Rate Trend", y = "%", x = "Year") +
  theme_minimal()

```

## Renewable Energy Share

```{r}

renewData <- allData[indicatorName == "Renewable energy consumption (% of total final energy consumption)"]
ggplot(renewData, aes(x = year, y = value, color = countryName)) +
  geom_line() +
  labs(title = "Renewable Energy Share", y = "%", x = "Year") +
  theme_minimal()

```

## Summary

-   Cleaned and merged data from 3 countries
-   Filtered, renamed, and explored indicators
-   Built reproducible plots and computations using R

Future work could include: - Modeling relationships between indicators - Forecasting trends using time-series models
