---
title: "Economic, Social, Environmental, Health, Education, Development and Energy of Albania, Croatia and Malta"
author: "Joel Thomas Chacko"
institute: "University College Dublin"
format:
  revealjs:
    theme: solarized
    slide-number: true
    incremental: true
    toc: true
    code-fold: true
    highlight-style: github
    center-title-slide: true
    transition: slide
editor: visual
---

## Introduction

This presentation analyzes Economic, Social, Environmental, Health, Education, Development, and Energy indicators for Albania, Croatia, and Malta using data from the World Bank via Humanitarian Data Exchange.

We use the `data.table` package in R for all required data manipulation tasks. Only where explicitly allowed, other packages are used (e.g., for visualization).

## Data Preparation

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
options(datatable.print.topn = 100)
```

## Load and Clean Data

```{r load-data}
# Load three country files
albania <- fread("datasets/indicators_alb_albania.csv")
croatia <- fread("datasets/indicators_hrv_croatia.csv")
malta <- fread("datasets/indicators_mlt_malta.csv")

# Drop first row from each dataset
albania = albania[-1]
croatia = croatia[-1]
malta = malta[-1]

# Combine and ensure correct types
allData <- rbindlist(list(albania, croatia, malta))
str(allData)
```

## Clean & Filter

```{r clean}
# Standardize column names (for compatibility)
setnames(allData, old = c("Country Name", "Country ISO3", "Year", "Indicator Name", "Indicator Code", "Value"),
                 new = c("countryName", "countryCode", "year", "indicatorName", "indicatorCode", "value"))


# Ensure correct column types
allData[, year := as.integer(year)]
allData[, value := as.numeric(value)]
```

## Merge Datasets (already done by rbindlist)

We merged data for Albania, Croatia, and Malta using `rbindlist()`.

## Quick Exploration

```{r exploration}
allData[, .N, by = .(countryName)]
allData[, .N, by = .(indicatorName)][1:10]
```
