---
title: "Using the CountryHealthAnalysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the CountryHealthAnalysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(CountryHealthAnalysis)
library(ggplot2)
```

# Introduction

This vignette demonstrates how to use the main functions of the `CountryHealthAnalysis` package to:

- Load cleaned indicator data from `.csv` files
- Use print and summary methods for inspection
- Visualize top or selected indicators using the plot method

# Reading Country Indicator Data

```{r read}
irlPath <- system.file("extdata", "indicators_irl.csv", package = "CountryHealthAnalysis")
mltPath <- system.file("extdata", "indicators_mlt.csv", package = "CountryHealthAnalysis")

irl <- readIndicators(irlPath)
mlt <- readIndicators(mltPath)
```

# Print Method Examples

### Default: Print summary for top 3 indicators

```{r print-default}
print(irl)
```

### Print with selected indicators

```{r print-custom}
print(irl, indicatorList = c("Net migration", "Urban population (% of total)"))
```

# Summary Method Examples

### Default: Top 7 indicators

```{r summary-default}
summary(mlt)
```

### Custom: Summary for selected indicators

```{r summary-custom}
summary(mlt, indicatorList = c("CO2 emissions (metric tons per capita)", "Population, total"))
```

### Change number of top indicators

```{r summary-topN}
summary(irl, topN = 10)
```

# Plot Method Examples

### Default: Plot top 3 indicators

```{r plot-default, fig.width = 10, fig.height = 5}
plot(irl)
```

### Plot selected indicators

```{r plot-custom, fig.width = 10, fig.height = 5}
plot(irl, indicatorList = c("Net migration", "School enrollment, primary (% gross)"))
```

### Plot top 5 indicators using faceting

```{r plot-facet, fig.width = 10, fig.height = 6}
plot(mlt, topN = 5, facet = TRUE)
```

# Closing Notes

All methods (`print`, `summary`, `plot`) use S3 dispatch and accept additional arguments such as `indicatorList` and `topN` for customized exploration.
