---
title: "STAT40830 - Homework 1"
author: "Joel Thomas Chacko"
format: 
  pdf:
    geometry: "top=1.5cm, bottom=1.2cm, left=1cm, right=1cm"
    fontsize: 11pt
editor: visual
execute:
    echo: false
    warning: false
    message: false
    fig-width: 7
    fig-height: 5
    fig-align: center
---

```{r}

library(dplyr)
library(knitr)
library(ggplot2)
data(CO2)

```

# Introduction

The dataset that's used is the $CO_2$ dataset present in the `datasets`. The $CO_2$ data set basically contains data from a study from an experiment on the cold tolerance of grass plants. It records the rate of CO₂ uptake (photosynthesis) by different plants by exposing varying concentrations of ambient $CO_2$, under different treatment conditions.

# Descriptive Statistics

```{r}

summaryTable <- CO2 |>
  group_by(Type, Treatment) |>
  summarise(
    Mean_Uptake = round(mean(uptake), 2),
    SD_Uptake = round(sd(uptake), 2),
    Count = n()
  )

kable(summaryTable, caption = "Mean Carbon Dioxide Uptake by Type and Treatment")
```

**Interpretation**

From the above table, we can see that there 21 plants in each group, and we observe the following

-    \textcolor{red}{Highest} $CO_2$ Absorption: **Quebec - Nonchilled plants**

-   \textcolor{blue}{Lowest} $CO_2$ Absorption: **Mississippi - Chilled plants**


