---
title: "Homework 1 - Adv Data Prog with R"
author: "Joel Thomas Chacko - 24220504"
format:
  html:
    fontsize: 1.2em
    toc: true
    toc-title: '**Contents**'
    embed-resources: true
    grid:
      body-width: 800px
  revealjs: 
    slide-number: true
    html-math-method: mathjax
    theme: serif
execute: 
  eval: true
  echo: false
  warning: false
  message: false
---

```{r}

library(dplyr)
library(knitr)
library(ggplot2)
data(CO2)

```

## Introduction

The dataset that's used is the $CO_2$ dataset present in the `datasets`. The $CO_2$ data set basically contains data from a study from an experiment on the cold tolerance of grass plants. It records the rate of CO₂ uptake (photosynthesis) by different plants by exposing varying concentrations of ambient $CO_2$, under different treatment conditions.

## Descriptive Statistics

```{r}

summaryTable <- CO2 |>
  group_by(Type, Treatment) |>
  summarise(
    Mean_Uptake = round(mean(uptake), 2),
    SD_Uptake = round(sd(uptake), 2),
    Count = n()
  )

kable(summaryTable, caption = "Mean Carbon Dioxide Uptake by Type and Treatment")
```

### Interpretation

From the above table, we can see that there 21 plants in each group, and we observe the following

-   \textcolor{red}{Highest} $CO_2$ Absorption: **Quebec - Nonchilled plants**

-   \textcolor{blue}{Lowest} $CO_2$ Absorption: **Mississippi - Chilled plants**

## Plot - Effect of Chilling on CO2 Uptake by Plant Type

```{r}


ggplot(CO2, aes(x = Treatment, y = uptake, fill = Type)) +
  geom_boxplot() +
  labs(
    title = "Effect of Chilling on CO2 Uptake by Plant Type",
    x = "Treatment Condition",
    y = "CO2 Uptake (μmol/m²/sec)",
    fill = "Plant Type"
  ) +
  theme_minimal()

```

### Interpretation

From the above plot we can clearly see that Quebec plants has a higher $CO_2$ update than the Missisippi plants, and we also see that that nonchilled plants (in comparatively less cold) environments have more absorption than plants in chilled environments.
